A Handbook for Tenured Track Faculty in Computer Science in the US

### 2024-07-08 faculty.tex
updates

Toward the end of your 3rd year (or early 4th year), you will need to prepare and submit a renewal package.  The department will then review your package and vote on whether to renew you for your 4th (and perhaps also 5th year). Many faculty members actually do not get renewed, so do take this process seriously.

@@

Both GMU and UNL ( and likely other universities) have a very strict and specific format for the CV, e.g., what to include, how to format, order of sections and subsections, etc.  
Fortunately I have always been using the university CV format and also kept my CV up-to-date, so I didn't have to do much. 
I did have to add more \emph{personal} details (e.g., teaching evaluation and funding amounts that match what the university has in their system).

@@

\subsection{Letters}
You will need letters from senior people in your area.  Someone who can vouch for you.
You will need 4--6 letters. However, you will not be the one asking for these letters, but your department will. You can suggest who to ask.  This process doesn't take much of your time (because you don't write the letters) but it might cause you a lot of stress (because you don't have full control over it).

@@

In my case I put just 2 names in my list (0 on blacklist) and let the department decide the rest. The reason that I only put 2 names is that I feel I don't have too many "enemies" out there that would write bad things about me.  Those that don't know me well will likely just decline the letter invitation.  This also gives the department more choices on their list.
Note I also thought about leaving my list empty, but then that seems to be too weird, and I don't want to give the impression that I don't have any "friends" in academic.

The department selected 1 of the 2 names on my list and 4 more names from the department's list. Moreover, it appears that no one declines so my letter writers are all set fairly quickly in the process. All of my letters were very good and give many positive comments about me as summarized in my evaluation letter (sent to me after the tenure decision).  I will never know who wrote the letters, but I am very grateful to them.

\subsection{Tenure Talk}
After you submit your tenure package, you might be asked to give a tenure talk.  This is a talk where you summarize your research, teaching, and services, and future plans.  The talk is typically 1 hour long, and there will be a Q\&A session afterward.  The talk is open to the public.

\subsection{Others}

>>>>

Toward the end of your 3rd year (or early 4th year), you will need to submit a \emph{renewal} application.  The department will then review your application and vote on whether to renew you for your 4th (and perhaps also 5th year). At some university this renewal is almost guaranteed, but at some other universities, it is not and people do get denied.

@@

For P\&T (Promotion and Tenure), universities often have a very strict and specific format for the CV, e.g., what to include, how to format, order of sections and subsections, etc.  
Fortunately I have always been using the university CV format and also kept my CV up-to-date, so I didn't have to do much (here's my most \href{https://dynaroars.github.io/people/nguyenthanhvuh/latex-cv/cv-nguyen.pdf}{recent CV}). 
I did have to add more \emph{personal} details (e.g., teaching evaluation and funding amounts that match what the university has in their system).

@@

\subsection{Letters}
You will need 4--6 letters from senior people in your area to vouch for your work. Your department will be ask people for these letters on your behalf. You can suggest who to ask.  This process doesn't take much of your time (because you don't write the letters) but it might cause you a lot of stress (because you don't have much control over it).

@@

I had just 2 names on my list (0 on blacklist) and let the department decide the rest. I feel I don't have too many ``enemies'' out there that would write bad things about me ( those that don't know me well will likely just decline the letter invitation instead of writing a bad letter).  This also gives the department more choices on their list (recall that people on the dept. list often weight more).
Note I also thought about leaving my list empty, but then that seems to be too weird. I don't want to give the impression that I don't have any ``friends'' in academic.

The department selected 1 of the 2 names on my list and 4 names from the department's list. It appears that no one declines, so my letter writers were all set fairly quickly. All of my letters were very good and give positive comments about me as summarized in my evaluation letter (sent to me after the tenure decision).  I will never know who wrote the letters, but I am very grateful to them.

\section{Others}
\subsection{Tenure Talk}
After you submit your tenure package, you might be asked to give a tenure talk, which summarizes your research, teaching, and services, and future plans.  The talk is open to the public, about 1 hour long, and has Q\&A session afterward.

Here is my \href{https://dynaroars.github.io/people/nguyenthanhvuh/files/pt/slides.pdf}{talk}. I focus on my recent research direction on neural network verification and just briefly on my past work. Overall, I didn't have to spend too much time on preparing because I reused slides from my job and research talk and just added at the end a ``stats'' slide on my funding, pubs, teaching, services, etc.

\subsection{Your ``Champion'' in the P\&T Meeting}

You will need a senior faculty member (or ``champion'') from your department to present your case in the department P\&T meeting. This person will present your achievements and argue for you.  Typically, this person is your mentor, but it can be anyone in the department that you trust and who knows your work well.
A good champion won't likely be able to help a bad case, but a bad champion can hurt a good case.  So do pay attention to this.

At GMU I was assigned a mentor, Paul Ammann, right when I joined.  For my tenure case, he put together my ``stats'' page and asked me to review it. I also discussed with him about letter writers, etc.  Overall, I think my case is relatively strong and so the whole thing was very ``smooth sail'' according to Paul. However, I did get much advice and guidance from him in other things e.g., how to run the MS SWE program (more on that later).

\begin{commentbox}[Vu:]
    While I have had not experienced with an ineffective mentor, I have heard about cases where the mentor did not do a good job, which can hurt strong candidates during the 4-th year renewal process. For example, a candidate might have few, but impactful high-quality publications, and the mentor did not emphasize this enough. In some case, an ineffective mentor actually turn a significant accomplishment into a negative, e.g., getting a 10-year Most Influential paper award is a big deal in certain fields, but the mentor did not emphasize this or allow it being downplayed by others (and allow idiotic comments from other faculty such as the work has been done 10 year ago). So do pay attention to this and make sure you have a good mentor who can defend you. If necessary, you can ask your chair to assign you a different mentor.

\end{commentbox}
\subsection{Others}
